---
title: "Week 3: Lab - Visualization "
author: 'Adam Saoud'
date: '1/20/2024'
output: html_document
---

---

# Instructions

In Chapter 12, we explored many different ways to “look at” the numbers. 
For this lab, let’s explore the `mtcars` dataset that is included within R.

This activity description does not provide the same level of code prompts as 
previous labs – it is assumed that you remember or can look up the necessary code. 
The overall goal of this activity is to use `ggplot2` to show different attributes 
of the `mtcars` dataset. Please be sure to include both the code and the images 
that were generated with your assignment.

Add all of your libraries that you use for this assignment here.
```{r setup}

library(ggplot2)

```

# Generate the following visualizations:

## Step 1 - Histogram

Histogram of `mpg`

```{r, "Step 1"}

ggplot(mtcars, aes(mpg)) +
  geom_histogram()


```

---

## Step 2 - Boxplots

Boxplots of `mpg` by `cyl` (i.e. 3 box plots: one for all cars with 4 cylinders, 
one for all cars with 6 cylinders, and one with all the cars with 8 cylinders).

```{r, "Step 2"}

ggplot(mtcars, aes(factor(cyl), mpg)) +
  labs(x = "Number of Cylinders", y = "Miles per Gallon") +
  geom_boxplot()

```

---

## Step 3 - MultiLine chart 

MultiLine chart of `wt` on the x-axis, `mpg` for the y-axis. 
With a line for each `am` (i.e. two lines). Also be sure to show each point on the chart.

```{r, "Step 3"}

ggplot(mtcars, aes(x = wt, y = mpg, color = factor(am))) + 
  labs(x = "Weight", y = "Miles per Gallon") +
  geom_point() + 
  geom_line()

```

---

## Step 4 - Barchart

Barchart with the x-axis being the name of each car, and the height being `wt`.
Make sure to rotate the x-axis labels, so we can actually read the car name.

```{r, "Step 4"}

ggplot(mtcars, aes(x = rownames(mtcars), y = wt)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  xlab("Car") +
  ylab("Weight") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

---

## Step 5 - Scatter chart
Scatter chart with the x-axis being the `mpg` and the y-axis being the `wt` of the car. 
Have the color and the size of each “symbol” (i.e., circle) represent how fast 
the car goes (based on the `qsec` attribute).

```{r, "Step 5"}

ggplot(mtcars, aes(x = mpg, y = wt, color = qsec, size = qsec)) +
  geom_point() +
  labs(x = "Miles per Gallon", y = "Weight")
  scale_size(range = c(1, 4))

```